### 2024-04-23 manuscript/main.Rmd

---
documentclass: article
classoption: [a4paper,11pt,twocolumn]
subparagraph: yes
header-includes:
  - \usepackage[utf8]{inputenc}
  - \usepackage[T1]{fontenc}
  - \usepackage{tipa}
  #- \usepackage[leqno,fleqn]{amsmath}
output:
  pdf_document:
    highlight: zenburn
    includes:
      before_body: includes/tex/doc_prefix.tex
      in_header: includes/tex/header.tex
    keep_tex: yes
    number_sections: yes
bibliography: includes/bib/icphs2023.bib
---

```{r, 'setup', include=FALSE}
library(knitr)
f <- function(x, options) {
  lab <- paste0(options$fig.lp, options$label)
  paste(
        "\\begin{figure}[!h]\n\\caption{", options$fig.cap.top, "}\\label{", lab,"}\n", 
        hook_plot_tex(x, options), 
        "\\end{figure}\n\n", sep = "")
}
knit_hooks$set(plot = f)
```

\section{Introduction}

Write something here

\section{Background Literature}

\subsection{The production of the alveolar fricative in Catalan and Spanish}

Both Catalan and Peninsular Spanish sound inventories have an phoneme /s/, which can be classified as a voiceless alveolar fricative \cite{hualde_colina_2015}. In addition, most of the Catalan varieties have a voiced alveolar phoneme in intervocalic contexts **(Lloret, 2011; Recasens Vives, 1996, p. 279)**. In onset position, /s/ and /z/ contrast phonemically, which leads to minimal pairs like *cel* [s\textepsilon l] - *zel* [z\textepsilon l] and *fossa* [fósa] - *fosa* [fóza] **(Carrera-Sabaté & Prieto, 2012-2019)**. Conversely, at the end of the word, this contrast is neutralized, being the voicing or devoicing of the alveolar fricative dependent on the following sound **(Institut d’Estudis Catalans, 2022)**. For example, if it is followed by a vowel, the alveolar fricative is voiced: *torres* [tór\textschwa{}s] and *torres antigues* [tor\textschwa{}z\textschwa{}nti\textgamma{}\textschwa{}s].

Interestingly, Medieval Spanish, similarly to Modern Catalan, distinguished between /s/ and /z/ in intervocalic position within a word, although these sounds merged through a devoicing process \cite{hualde_colina_2015}. Nevertheless, in those modern varieties where the /s/ in coda position is produced, such as in Peninsular Spanish, this voiceless phoneme might assimilate to the following sound, especially if it is a voiced consonant (*mismo*, [mízmo]), rather than a vowel (*las alas*, [lasálas]) **(Hualde & Colina, 2015; Hualde et al., 2020)**. Hence, Spanish inventory does not have the phoneme /z/, but [z] may occur as an allophone of /s/, due to a voice assimilation process. It is important to mention, though, that it has been claimed that this voicing is generally not complete, but partial **(Benet et al., 2012; Hualde & Colina, 2015; Hualde et al., 2020)**.

There is some evidence of the influence of Spanish on the production of the Catalan alveolar fricative, in contexts where, according to prescriptivist rules **(Institut d'Estudis Catalans, 2022)**, the [z] should be produced, but [s] is produced instead. This phenomenon has mainly been attested in those areas where there has been an extended contact between Catalan and Spanish, such asthe so-called "franja de Aragón" (borderland between Catalonia and Aragon), some areas of the Valencian Community, and Barcelona and other urban areas **(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p. 268)**. The opposite pattern has also been observed - that is, the voicing of the alveolar fricative in the Spanish spoken in Barcelona and Valencia. Interestingly, although in both Barcelona and Valencia Catalan and Spanish coexist and influence each other, it seems that Catalan has a greater influence on Spanish in Barcelona, but  Spanish has a greater influence on Catalan in Valencia **(Davidson, 2020)**.

In summary, both Catalan and Spanish have a voiceless alveolar fricative in their inventories. Catalan also has a voiced variant, which is produced when this sound is followed by a voiced consonant or a vowel. Spanish might produced this sound only when it is followed by a voiced consonant, although /s/ might be "partially voiced" (Hualde & Colina, 2015). Thus, these phonemes, although they have the same allophones, their contextual distribution is different (Llisterri, 2023). Moreover, some studies have shown that there is a bidirectional influence between both languages in the production of the voiced vs the voiceless alveolar fricative, and that such influence might be different based on the context. These divergent patterns might be explained by the different sociopolitic and linguistic realities of each city, and  might have an effect on speakers' attitudes towards the devoicing of the alveolar fricative /z/ in Catalan.

\subsection{Language attitudes towards the alveolar fricative in Catalan and Spanish}

The social meaning and the perception of intervocalic [z] in Spanish has been studied in a variety of contexts, including both monolingual **(Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell & García, 2017; García, 2015, 2019; Silverstein, 2003)** and bilingual contexts **(Davidson, 2014, 2015, 2019; Sinner, 2002)**, despite having been categorized as a low-salient variant **(Chappell, 2017, 2021; Chappell et al., 2023; García, 2015)**. With respect to the former, it has been found that for listeners from San José, Costa Rica, the covert attitudes associated with intervocalic [z] vary significantly by speaker gender. Consequently, male speakers who use [z] tend to be positively evaluated, and generally perceived as nice, local, confident, and masculine. Conversely, female speakers who use [z] are only afforded negative associations of lower education and social class **(Chappell, 2016; Chappell et al., 2023; Silverstein, 2003)**.

Similarly, **García (2019)** shows that, in Ecuador, intervocalic /s/ is evaluated differently in male and female speakers. Accordingly, women who produce intervocalic [z] are viewed as significantly lower status, less pleasant, and older. Contrary, men who produce intervocalic [z] do not receive these negative evaluations. Additionally, this variable has been found to serve as a marker of regional identity, with voicing being associated with neighboring Highland Ecuadorian dialects (Cuenca and Quito) that exhibit systematic voicing. **García (2020)** hypothesizes that young men are the leaders of this change, essentially due to the fact that they are in greater contact with these highland varieties. Interestingly, the mean voicing rates of females are roughly two age groups behind those of males. Consequently, it is believed that it is not necessarily the case that women in Ecuador are avoiding voicing like San José women due to its negative evaluations, but rather that they are further behind in the adoption of the incoming variant.

With respect to the social meaning and the perception of intervocalic [z] in bilingual contexts, this phenomenon has been mainly studied in Catalan-Spanish contact situations, given its robust usage by Catalan-Spanish bilinguals and, in particular, by those with greatest exposure to and usage of Catalan **(Davidson, 2014, 2015)**. According to **Sinner (2002)**, intervocalic [z] can be considered a possible linguistic marker of Catalonian Spanish, despite its relative lack of social awareness **(Davidson, 2019; Sinner, 2002)**. In their study, Sinner (2002) interviewed 12 speakers of Barcelonan Spanish and monolingual Spanish-speakers from Madrid aged 27 to 41 regarding their awareness of linguistic features of Catalanized Spanish. Results showed that only select Barcelona speakers perceived intervocalic [z] as a linguistic marker of Catalonian Spanish, whereas monolingual Spanish-speakers from Madrid did not. Additionally, Davidson (2019) found that, in Barcelona, the indexical field of intervocalic [z] features exclusively positive associations with bilingualism and local solidarity, which warrants its gradual adoption in the speech community, unlike other phonetic markers of Catalonian Spanish that are the topic of overt social ridicule, such as lateral velarization. These results are in line with previous findings on intervocalic [z] in Catalonian Spanish. For instance, **Silverstein (2003)** found that intervocalic [z] indexes a local Barcelona identity, and the variant then acquires positive associations with bilingualism and cultural pride. Overall, these findings suggest that intervocalic [z] likely functions as a sociolinguistic indicator of Catalonian Spanish within the Catalonian speech community, indexing Catalan-Spanish bilingual membership while lacking any considerable degree of generalized conscious awareness **(Davidson, 2019)**.

\subsubsection{Headings}

Section headings are centered in boldface with capitalized letters. Sub-headings start at the left margin in the column with the first letter capitalized and the rest of the heading in lower case. Sub-sub-headings appear like sub-headings, except that they are in italics and not boldface. See examples in this file. No more than 3 levels of headings should be used.  Empty lines should be left above and below each section heading.

\subsection{Text font}

Times or Times New Roman font is used for the main text. Recommended font size is 11 points. Other font types may be used if needed for special purposes. If using any non-Unicode fonts, these must be embedded in the final PDF file.

The \LaTeX{} template can be used either with plain \LaTeX{} or \XeLaTeX.

\subsection{Figures}

All figures should be centered on the column (or page, if the figure
spans both columns). Figure captions should follow each figure and
have the format given in Fig. \ref{fig:vowels}.

Figures should preferably be line drawings. If they contain grey
shades or colours, it should be checked that they print well on a
high-quality noncolour laser printer.

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=6cm]{./includes/figures/ipa.eps}
\caption{The vowel chart used in the International Phonetic
Alphabet (IPA).}\label{fig:vowels}
\end{center}
\end{figure}

\subsection{Tables}

An example of a table is shown as Table~\ref{tab:decibel}.  Somewhat
different styles are allowed according to the type and purpose of the
table. Colour should not be used, but grey shading is allowed. There
should be a margin of 6~points (pt) above and below the table.

The caption text may be above or below the table, but this should be
consistent throughout the submission. Left and right indentation of
the caption should be 0.5~cm.

\begin{table}[!ht]
  \begin{center}
  \begin{tabular}{|c|c|}
  \hline
  \rowcolor[gray]{.75}
  ratio    & Decibels\\
  \hline
  $1/1$    & $0$\\
  $2/1$    & $6$\\
  $3.16$   & $10$\\
  $1/10$   & $-20$\\
  $10/1$   & $20$\\
  $100/1$  & $40$\\
  $1000/1$ & $60$\\
  \hline
  \end{tabular}
  \caption{This is an example of a table showing Decibel (dB)
  ratios.}\label{tab:decibel}
  \end{center}
\end{table}

\subsection{Equations}

Equations should be placed on separate lines and numbered. An example
of an equation is
\begin{equation}\label{eq:tzero}
T_0 = \frac{1}{f_0}.
\end{equation}
Numbers of equations can be on the right or on the left margin of the
text column.

\subsection{Examples}

Examples from other languages can either be presented in the body text, or, if referred to elsewhere or particularly long and complex, can be put on a separate, numbered line, as should be done for equations.

\subsection{Phonetic fonts}

We recommend that you use the \textsc{tipa} package for IPA phonetic symbols. For information about how to access Unicode fonts in the Word template, see \cite{IPA-SIL} or \cite{IPA-KEYBOARD}. The font you use must be embedded. Please remember to check this,  e.g. by inspecting the ``Document Properties --- Fonts'' in Acrobat Reader.

\subsection{Page numbering}

Page numbers will be added electronically to the document
later. \textit{Please do not add page numbers and please do not insert
any footers or headers!}

\subsection{References}

Please use just the reference number in square brackets. Formulations
with author names like "\ldots\ as Ladefoged \cite{Ladefoged:2003}
showed that \ldots" are acceptable but not "as shown in [Ladefoged,
3]" or "as shown in (Ladefoged [3])".

References are to be numbered in alphabetical order. Please
double-check the final version of your paper with regard to the
correct correspondence of references to their numbers.

\subsection{Hyperlinks}

Links to URLs or email addresses should be formatted as normal text,
\textit{not} as hyperlinks and not blue or underlined etc. Usually
hyperlinks to web pages are listed in the references section. If
required, line breaks can be placed within URLs after slashes or
dashes (cf.~\cite{IPA-SIL,IPA-KEYBOARD}), but doublecheck that no hyphens
are inserted.

\subsection{Footnotes and endnotes}

If footnotes cannot be avoided they should appear as
endnotes.\footnote{This footnote appears here as an endnote.}

\section{Multimedia files}

Multimedia data that are part of the paper are to be embedded in the submitted PDF; they cannot be submitted as supplementary data. Any images are to be included in the paper as Figures (see Section 2.3 above). It is the authors' responsibility to check image quality ahead of submission. Audio examples are to be embedded within the PDF. To do this, authors can generate the PDF, and then embed the audio files using Adobe Acrobat Professional, or other software that offers the same outcome, so that the audio is included in the PDF. The presence of audio data should be identified in the text.

We encourage authors to illustrate video data using still photographs from the video, and to include them as figures in the PDF.  We cannot accept embedded video files, but authors are welcome to refer readers to a URL on the internet where these can be viewed.

\section{PDF details}

PDF files submitted must comply with the following requirements:

\begin{enumerate}
\item all special fonts and symbols must be embedded in the PDF
  file so that correct rendering of the PDF does not depend on the
  fonts installed on the viewer's computer
\item there must be no
  password protection on the PDF file, i.e. PDF files must not be
  protected by PDF security in any way, i.e. content extraction,
  document assembly, high resolution printing etc. must not be
  forbidden
\item PDF files should not contain any colours, hyperlinks,
  multimedia or 3D content, and no JavaScript or forms
\item PDF files
  should be no larger than 5 MB.
\end{enumerate}

\section{Anonymity}

In ICPhS 2023 submissions, an anonymous reviewing process will be used. This means that for the first submission the name(s) of the author(s) and
their affiliation(s) \emph{must not} be mentioned. In addition, please refrain from using acknowledgements.  Please also try to make your own previous research as anonymous as possible. As an example: do not write "In our previous study \cite{Stevens:1999} we could show\ldots" but "As shown in \cite{Stevens:1999}\ldots". Or refer to your own published or otherwise widely known work, and to that of the other authors, in the "Julius Caesar style", i.e. in the third person (for example: his work, her work, their work). Reference as "anonymous" only work that you or the other authors have submitted for publication, but which has not yet been published, e.g. \cite{Anonymous:submitted}.

Please make sure that no author details appear in the Document Properties of the PDF file. \emph{For the revised paper submission author details are of course needed}. Acknowledgements and references to one's own work are possible as usual.

\section{Format of references}

References are formatted using the IEEE standard (available in various reference management systems like Zotero or Mendeley). If you do not use a reference management system, please use the examples provided for monographs \cite{Fant:1960}, contributions to volumes \cite{Stevens:1999}, journal articles \cite{Beattie/etal:1982}, articles in conference proceedings \cite{Ladefoged:2003}. Abbreviations of
well-known conferences and journals are possible \cite{Peterson/Barney:1952}. Software tools \cite{Boril/Skarnitzl:2016} are referenced according to authors' instructions.

\section{R scripts}

You can use knitr code chunks like in any .Rmd document:

```{r, 'chunk_ex'}
2 + 2
```

This includes inline code: 2 + 2 = `r 2 + 2`.

R scripts can be loaded as well:

```{r, 'load_script', echo=TRUE, message=FALSE, warning=FALSE}
source("./includes/scripts/analysis.R")
```

Which means you have access to any objects assigned in the script, like Figure 
\ref{fig:ex_plot}:

```{r, 'ex_plot', fig.align='center', fig.width=2.75, fig.height=1.5, fig.cap.top="This is a figure caption.", fig.pos='h', fig.lp='fig:'} 
cars_plot
```

\bibliographystyle{./includes/bib/IEEEtran.bst}
\bibliography{./includes/bib/icphs2023.bib}

\theendnotes