### 2024-05-03 manuscript/main.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  a4paper,
  11pt,
  twocolumn]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
%------------------------------------------------------------------------------%
% PAPER TEMPLATE FOR ICPHS 2023 Prague                                         %
%                                                                              %
% Original template downloaded from:                                           %
% http://www.icphs2023.org/call-for-papers/                                    %
%                                                                              %
% Reformatted to work with Rmarkdown and R by:                                 %
% Joseph V. Casillas | Rutgers Univesity |11/11/2022                           %
%                                                                              %
% Available for download at:                                                   %
% https://github.com/jvcasillas/icphs2023_rmd_template                         %
%------------------------------------------------------------------------------%

% Packages
\usepackage{./includes/tex/icphs2023}
\usepackage{metalogo} 
\usepackage{epstopdf}
\usepackage{tipa}

% Links and urls must be black
%\hypersetup{urlcolor=black, citecolor=black, linkcolor=black}

% Packages removed from icphs2023.sty because of conflicts
% They have been added to the .Rmd yaml front matter
% \usepackage[latin1]{inputenc}
% \usepackage[T1]{fontenc}
% \usepackage[leqno,fleqn]{amsmath}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tipa}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\title{Language Attitudes of Catalan-Spanish Bilinguals Towards Intervocalic Alveolar Fricatives in Catalan}
\author{Please write XXXX instead of the name(s) of the author(s)}
\organization{Please write XXXX instead of the affiliation(s)}
\email{please write XXXX instead of the email address(es)}

\maketitle

\begin{abstract}
Catalan and Spanish are two languages that have coexisted for centuries in Eastern Spain. Thus, there is a bidirectional influence between these languages in all linguistic levels, including in the production of sounds. However, little is known about the perception and the attitudes toward a less Catalan-like production of a given sound. Therefore, the aim of this study is to examine the covert attitudes toward the production of intervocalic /s/, as either [s] or [z] by Catalan-Spanish bilinguals from Catalonia using a matched guise test. Results indicate that phoneme type is not a significant predictor of language attitudes in Catalonia, but other individual social factors, such as mother tongue and province, might be.
\end{abstract}

\keywords{Catalan, Spanish, language attitudes, intervocalic alveolar fricatives.}

\section{Introduction}

Catalan and Spanish are two languages that have coexisted for centuries,
mainly in eastern Spain, along the Mediterranean coast
\cite{hualde_1992} It is safe to assume that there is a bidirectional
influence between these languages: in that region, Catalan has
influenced Spanish and Spanish has influenced Catalan in all linguistic
levels \cite{davidson_2014} \cite{davidson_2015}. A crucial area of
cross-linguistic influence might be the production of sounds, especially
in those sounds that have different distribution in Catalan and Spanish.
For example, previous studies have found that, in Barcelona, the
production of Catalan vowels is different based on the L1 of the
speaker, being it less target-like in L1 Spanish speakers compared to L1
Catalan speakers \cite{pallier_bosch_sebastian_galles_1997}.

However, little is known about the perception and the attitudes toward a
less Catalan-like production of a given sound. Therefore, the aim of
this study is to examine the covert attitudes towards the production of
intervocalic /s/, as either {[}s{]} (disfavored by prescriptivist norms)
or {[}z{]} (preferred by prescriptivist norms) by L1 Catalan, L1
Spanish, and Catalan-Spanish bilinguals from Catalonia using a matched
guise test \cite{lambert_1967}. The alveolar fricative was chosen
because it exists in both Catalan and Spanish inventories, but their
contextual distribution is different \cite{llisterri_2023}.

\section{Background Literature}

\subsection{The production of the alveolar fricative in Catalan and Spanish}

Both Catalan and Peninsular Spanish sound inventories have a phoneme
/s/, which can be classified as a voiceless alveolar fricative
\cite{hualde_2015}. In addition, most of the Catalan varieties have a
voiced alveolar phoneme in intervocalic contexts \cite{lloret_2011}
\cite{recasens_1996}. In onset position, /s/ and /z/ contrast
phonemically, which leads to minimal pairs like \emph{cel}
{[}s\textepsilon l{]} - \emph{zel} {[}z\textepsilon l{]} and
\emph{fossa} {[}fósa{]} - \emph{fosa} {[}fóza{]}
\cite{carrera_sabate_prieto_2012_2019}. Conversely, at the end of the
word, this contrast is neutralized, being the voicing or devoicing of
the alveolar fricative dependent on the following sound \cite{iec_2022}.
For example, if it is followed by a vowel, the alveolar fricative is
voiced: \emph{torres} {[}tór\textschwa{}s{]} and \emph{torres antigues}
{[}tor\textschwa{}z\textschwa{}nti\textgamma{}\textschwa{}s{]}.

While Spanish inventory does not have the phoneme /z/, the sound {[}z{]}
may occur as an allophone of /s/, due to a voice assimilation process.
It is important to mention, though, that it has been claimed that this
voicing is generally not complete, but partial
\cite{benet_cortes_lleo_2012} \cite{hualde_2015}
\cite{hualde_et_al_2020}. There is some evidence of the influence of
Spanish on the production of the Catalan alveolar fricative, in contexts
where, according to prescriptivist rules \cite{iec_2022}, the {[}z{]}
should be produced, but {[}s{]} is produced instead. This phenomenon has
mainly been attested in those areas where there has been an extended
contact between Catalan and Spanish, such as the so-called ``franja de
Aragón'' (borderland between Catalonia and Aragon), some areas of the
Valencian Community, and Barcelona and other urban areas
\cite{benet_cortes_lleo_2012} \cite{recasens_1996} \cite{davidson_2020}.
The opposite pattern has also been observed - that is, the voicing of
the alveolar fricative in the Spanish spoken in Barcelona and Valencia.
Interestingly, although in both Barcelona and Valencia Catalan and
Spanish coexist and influence each other, it seems that Catalan has a
greater influence on Spanish in Barcelona, but Spanish has a greater
influence on Catalan in Valencia \cite{davidson_2020}.

\subsection{Language attitudes towards the alveolar fricative in Catalan and Spanish}

The social meaning and the perception of intervocalic {[}z{]} in Spanish
has been studied in a variety of contexts, including both monolingual
\cite{chappel_et_al_2023} \cite{chappel_2021} \cite{garcia_2019}
\cite{chappel_2016} \cite{chappell_2017} \cite{chappell_garcia_2017}
\cite{silverstein_2003} and bilingual contexts \cite{davidson_2014}
\cite{davidson_2015} \cite{davidson_2019} \cite{sinner_2002}, despite
having been categorized as a low-salient variant
\cite{chappel_et_al_2023} \cite{chappel_2021} \cite{chappell_2017} .
With respect to the former, it has been found that for listeners from
Costa Rica and Ecuador \cite{garcia_2019}, the covert attitudes
associated with intervocalic {[}z{]} vary significantly by speaker
gender. Consequently, male speakers who use {[}z{]} tend to be
positively evaluated, and generally perceived as nice, local, confident,
and masculine. Conversely, female speakers who use {[}z{]} are only
afforded negative associations of lower education and social class
\cite{chappel_et_al_2023} \cite{chappel_2016} \cite{silverstein_2003}.
Additionally, this variable has been found to serve as a marker of
regional identity, with voicing being associated with neighboring
Highland Ecuadorian dialects that exhibit systematic voicing.

With respect to the social meaning and the perception of intervocalic
{[}z{]} in bilingual contexts, this phenomenon has been mainly studied
in Catalan-Spanish contact situations \cite{davidson_2014}
\cite{davidson_2015}. According to Sinner \cite{sinner_2002},
intervocalic {[}z{]} can be considered a possible linguistic marker of
Catalonian Spanish, despite its relative lack of social awareness
\cite{davidson_2019} \cite{sinner_2002}. Additionally, Davidson (2019)
found that, in Barcelona, the indexical field of intervocalic {[}z{]}
features exclusively positive associations with bilingualism and local
solidarity, which warrants its gradual adoption in the speech community,
unlike other phonetic markers of Catalonian Spanish that are the topic
of overt social ridicule, such as lateral velarization. These results
are in line with previous findings on intervocalic {[}z{]} in Catalonian
Spanish. For instance, Silverstein \cite{silverstein_2003} found that
intervocalic {[}z{]} indexes a local Barcelona identity, and the variant
then acquires positive associations with bilingualism and cultural
pride. Overall, these findings suggest that intervocalic {[}z{]} likely
functions as a sociolinguistic indicator of Catalonian Spanish within
the Catalonian speech community, indexing Catalan-Spanish bilingual
membership while lacking any considerable degree of generalized
conscious awareness \cite{davidson_2019}.

\subsection{The present study}

Accordingly, and considering that there are no perception studies
focusing on Catalonian Catalan (central variety), the present study aims
to answer the following research questions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  What covert attitudes are linked to the perception Catalan
  intervocalic /s/ as either {[}s{]} or {[}z{]}?
\item
  What is the relationship between individual social factors and covert
  attitudes towards Catalan intervocalic /s/?
\end{enumerate}

\section{Methods}

\subsection{Materials}

Data were collected using an online survey hosted on Qualtrics that
consisted of two parts. The first part, a demographic questionnaire,
collected information regarding participant age, gender, place of
origin, place of residence, mother tongue, family income, and level of
studies. The second portion of the survey consisted of a perception task
where participants were presented with a series of stimuli containing
the target phoneme contrast (intervocalic voiced and voiceless alveolar
fricative) such as \emph{Han tancat les vies estretes} `they closed the
narrow streets', followed by 11 evaluative statements related to social
characteristics such as level of studies, friendliness, social class,
and ethnic origin. These evaluative statements were then followed by a
100-point draggable sliding scale with which participants indicated
their level of agreement or disagreement with the content of the
statement, where 0 was ``totally disagree'' and 100 was ``totally
agree''.

The target stimuli were recorded by a 28-year-old female native speaker
of Catalan in a whisper room 6084 E sound booth using a personal
communication device. The stimuli were between 3 and 4 seconds long and
were controlled for utterance length (5 words), syllable count (2
syllables for the target word and 3 for the following adjective), and
lexical stress (target sound in unstressed position). A total of 12
target stimuli containing intervocalic voiced /s/ were recorded, as well
as different samples of the voiceless /s/. The target sentences were
manipulated in Praat by splicing the voiced /s/ and pasting the
voiceless /s/ in order to create 2 conditions: 12 stimuli containing
intervocalic voiced /s/ and 12 stimuli containing intervocalic voiceless
/s/, for a total of 24 experimental sentences. These sentences were
combined with 32 fillers targeting other phonemic contrasts, for a total
of 56 sentences randomly distributed in two different versions. Each
included 12 experimental stimuli (6 x 2 conditions) and 16 fillers (8 x
2 conditions).

\subsection{Procedure}

Participants self-identified for this study. These were the
requirements: to be over 18 years of age, to be from Catalonia, and to
be proficient in both Catalan and Spanish. The survey was shared using
social networks and online communication channels and it followed a
between-subjects design. Therefore, participants were randomly assigned
to one of the two versions.

\subsection{Participants}

A total of 78 responses to the questionnaire were collected. Of these,
56 responses were incomplete and were therefore discarded. Data are
drawn from a total of 22 participants (13 females) aged between 18 and
72 years old (mean = 38) from Barcelona and Girona.

\subsection{Data analysis}

Data were analyzed in R (R Core Team, 2023). For each of the 11
evaluative statements, we created a series of hierarchically nested
linear regression models using the lme4 package (Bates et al., 2015)
with attitudinal sliding scale responses as the dependent variable, and
phoneme, mother tongue, and province as potential predictors. To
determine the best fit model for each statement, we used the anova()
function (Fox \& Weisberg, 2019).

\section{Results}

The statistical analysis did not reveal a significant effect of phoneme
on the evaluation of the speaker (p \textgreater{} .05) for any of the
11 evaluative statements. However, the model output did show a
significant association between mother tongue and province for some of
the statements. In what follows, we will explore the results for those
statements.

Participant responses to the statements \emph{Parla bé català} `S/he
speaks proper Catalan' and \emph{És d'origen català} `S/he is of Catalan
origin' show similar results, as illustrated in Figure \ref{fig:oc}.
Those who have both Catalan and Spanish as their L1 rate the speaker
more positively regardless of phoneme, compared to those whose L1 is
either Catalan or Spanish. In addition, although it is not statistically
significant, participants give higher ratings to the speaker when they
are listening to those statements with {[}z{]}.

\begin{figure}[!ht]
\begin{center}
\includegraphics[height=15cm]{./includes/figures/oc.png}
\caption{Ratings of "És d'origen català" as a function of mother tongue}\label{fig:oc}
\end{center}
\end{figure}

Participant responses to the statements \emph{És simpàtic} `S/he is
nice' and \emph{És de fiar} `S/he is trustworthy' showed similar
results, as illustrated in Figures \ref{fig:fiar}. Those who are from
and live in the province of Barcelona give higher ratings to the
speaker, compared to those who are from and live in Girona.

\begin{figure}[!ht]
\begin{center}
\includegraphics[height=15cm]{./includes/figures/fiar.png}
\caption{Ratings of "És de fiar" as a function of mother tongue}\label{fig:fiar}
\end{center}
\end{figure}

\section{Discussion}

The aim of this exploratory study was to explore the covert attitudes
that are linked to the perception Catalan intervocalic /s/ as either
{[}s{]} or {[}z{]} and to examine the relationship between individual
social factors and covert attitudes towards Catalan intervocalic /s/.
The results suggest that these predictors have a different effect on the
ratings of the speaker depending on the social characteristic under
evaluation.

Contrary to what we expected, phoneme was not a significant predictor.
One explanation for this divergent pattern could be that either
participants could not distinguish between {[}s{]} and {[}z{]}
\cite{davidson_2019} or, if they could, this difference in production
was not changing their attitudes towards the speaker \cite{sinner_2002}.
This finding is interesting considering that the production of the
voiceless alveolar fricative in intervocalic position has been
considered a result of Catalan in contact with Spanish
\cite{benet_cortes_lleo_2012} \cite{recasens_1996} \cite{davidson_2020},
and, from a prescriptivist point of view, it should be avoided
\cite{iec_2022}. Despite the lack of significance of this predictor, it
should be mentioned that the speaker received more positive ratings when
the {[}z{]} was being produced. The fact that phoneme was not an
important factor contrasts with Davidson \cite{davidson_2020}, who
analyzed covert and overt attitudes towards Catalonian Spanish and found
that {[}z{]} was a significant predictor: it was considered a
sociolinguistic indicator of Catalan-Spanish bilingual membership.
Although at this point we can only hypothesize, it could be that {[}z{]}
is perceptually more salient in Spanish than in Catalan, as the Spanish
inventory does not have the phoneme /z/, but voicing of /s/ might
partially occur \cite{benet_cortes_lleo_2012} \cite{hualde_2015}
\cite{hualde_et_al_2020}. Another explanation could be that the /s/
alone does not provide enough social information; rather, the
combination of the phoneme /s/ with the precedent vowel might give the
listener more cues and have a stronger social meaning. Future studies
should address this issue in order to explore if the combination of
vowel + /s/ carries other social connotations than just the /s/ alone.

We expected mother tongue to be a strong predictor of the attitudes
towards the production of intervocalic /s/, specifically, that L1
Catalan and bilingual speakers would give higher rates to those
statements with {[}z{]}. This hypothesis was only partially confirmed.
As reported above, phoneme was not a factor influencing the attitudes,
and mother tongue was only a significant predictor in the two statements
related to Catalan identity, namely \emph{Parla bé català} `S/he speaks
proper Catalan' and \emph{És d'origen català} `S/he is of Catalan
origin'. Interestingly, bilinguals gave higher ratings to the speaker,
followed by L1 Catalan participants and L1 Spanish participants. A
potential explanation for this finding, that should be verified in
future research, could be that L1 Catalan speakers have specific ideas
(higher standards) towards the `proper' Catalan. In addition, some
studies have suggested that those Catalan people who grow up in
Spanish-speaking households typically exhibit more negative attitudes
towards Catalan \cite{aguilo_more_lynch_2017}, which is consistent with
our results.

Lastly, province was also found as an important predictor, but only for
some statements. In two of the solidarity statements (\emph{És simpàtic}
`S/he is nice'; \emph{És de fiar} `S/he is trustworthy'), the speaker
was rated more positively by those participants from the province of
Barcelona. This finding might suggest that they identify with the
speaker; thus, giving her higher ratings \textbf{BUSCAR CITA}.
Interestingly, the speaker that created the stimuli is from the province
of Girona, but has spent the last 11 years of her life living outside
that area, either in Barcelona or in another country. We asked her to
read the sentences using a Central Catalan `neutral accent', which
people might have potentially related with Barcelona. This `neutral
accent' does not show any of the prototypical characteristics of the
\emph{accent gironí}, a subvariety within Central Catalan spoken in some
areas of Girona \cite{gros_2024}.

\section{Conclusion}

In summary, our findings suggest that the relationship between covert
attitudes and speaker ratings is not straightforward, as the speaker was
evaluated differently based on the social characteristic, with varying
factors influencing such ratings (mother tongue and province).
Crucially, phoneme production did not had an effect on the ratings, even
though prescriptivist norms disfavor the realization of {[}s{]} in
intervocalic position. Future studies should include a larger
participant pool, as well as participants from other regions in
Catalonia, such as Lleida and Tarragona. In addition, examining other
phoneme productions that emerge through the contact of Catalan with
Spanish could give us a better understanding of the factors that affect
the attitudes towards the Catalan language, and potential
sociolinguistic stereotypes and indicators related to Catalan identity.

\bibliographystyle{./includes/bib/IEEEtran.bst}
\bibliography{./includes/bib/icphs2023.bib}

\end{document}